ProbabilityBoundsAnalysis.jl:\\ Arithmetic with sets of distributions

### 2021-10-04 paper/header.tex
add why Julia section

\title{ProbabilityBoundsAnalysis.jl:\\ an arithmetic of probabilities and sets of probabilities}

@@

\hypersetup{
pdftitle = {ProbabilityBoundsAnalysis.jl:\\ an arithmetic of probabilities and sets of probabilities},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Ander Gray, Scott Ferson, Edoardo Patelli},
pdfkeywords = {Julia, Probability, Uncertainty, Interval Arithmetic, Probabilistic Arithmetic},
}

>>>>

\title{ProbabilityBoundsAnalysis.jl:\\ Arithmetic with sets of distributions}

@@

\hypersetup{
pdftitle = {ProbabilityBoundsAnalysis.jl:\\ an Arithmetic of probabilities and sets of probabilities},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Ander Gray, Scott Ferson, Edoardo Patelli},
pdfkeywords = {Julia, Uncertainty, Interval Arithmetic, Probability Bounds Analysis, Probability Boxes},
}

### 2021-10-04 paper/paper.tex
add why Julia section

\iffalse
\subsection{Automatic uncertainty in a programming language: why Julia?}

The long term goal of probability bounds analysis is to create a fully uncertain programming language, where any computer variable may be represented as an interval, distribution, p-box or other imprecise quantity. Such a framework would allow for uncertain extensions of deterministic functions to be computed in an automatic, rigorous and tight fashion. Those with little knowledge of probability theory can add uncertainty to their computations with minimal effort. Stochastic calculations often require many years of training to implement and interpret correctly, and are usually not rigorous and are famously computationally expensive. \texttt{ProbabilityBoundsAnalysis.jl} provides the type structure for distribution functions and p-boxes which is rigorous, and has many of the base operations required for an uncertain programming language, which are rigorous and computationally effient. In this section we argue why Julia is an ideal target language for such an uncertainty language.

@@

Multiple dispatch allows Julia packages with completely different functionalities to work together with minimal effort. One of the great successes of \texttt{IntervalArithmetic.jl} is that it allows for intervals to be propagated through nearly any Julia function, including those from different packages.

@@

Parametric typing allows one to create very generic data types, parametrised on other types. This is

For example, an uncertain

an interval of type Int64 (\texttt(Interval{Int64})) can be made and destiguised of t

@@

Blah blah blah

\subsubsection{Native parallelisation} \hfill \break

Blah blah blah

%, and discuss some of the remaining theoretical tasks required to make such a goal a reality. 
\fi
\section{Probability Boxes}
\label{sec:pboxes}

@@

@@

\section*{Acknowledgements}
We would like to thank our friends and colleagues at the Risk Institute who have supported \texttt{ProbabilityBoundsAnalysis.jl} from its birth, and through their contributions and discussions have made this software possible: Marco De Angelis, Nick Gray, Alex Wimbush and Dominic Calleja. We also thank Dominic and Liam Henry's (Clarus Financial Technology, Belfast) detailed review of this paper.

>>>>

\subsection{Automatic uncertainty in a programming language: why Julia?}

The long term goal of probability bounds analysis is to create a fully uncertain programming language, where any computer variable may be represented as an interval, distribution, p-box or other imprecise quantity. Such a framework would allow for uncertain extensions of deterministic functions to be computed in an automatic, rigorous and tight fashion. Those with little knowledge of probability theory could add uncertainty to their computations with minimal effort. Stochastic calculations often require many years of training to implement and interpret correctly, and are usually not rigorous and are famously computationally expensive. \texttt{ProbabilityBoundsAnalysis.jl} provides the arithmetic basis for such a framework. In this section we give some arguments why Julia \cite{bezanson2017julia} is an ideal target language for such an uncertainty language.

@@

Multiple dispatch allows Julia packages with completely different functionalities to work together with minimal effort. One of the great successes of \texttt{IntervalArithmetic.jl} is that it allows for intervals to be propagated through nearly any Julia function, including those from different packages. The same paradigm allows \texttt{ProbabilityBoundsAnalysis.jl} to propagate distributions and p-boxes through general Julia functions.

@@

Parametric typing allows one to create very generic data types, parametrised on other types. For example, an uncertain type of \texttt{Int64} (corresponding to an uncertain collection of finitely many numbers) could be distinguished to an uncertain type of \texttt{Float64} (an uncertain collection of reals). This is important because the behaviour and distribution of an uncertain collection of integers is different to that of real numbers. Parametric typing could generally be used to create an uncertain version of any Julia type, including custom types from other packages.

@@

An uncertain programming language would require \textit{source-to-source} transformations at times. For example, one way to evaluate a programs control-flow (\texttt{if-else} conditions) with uncertain inputs is to unravel the control flow, and to evaluate each possible conditional branch uniquely, and then combine the output. For example consider following function

\begin{lstlisting}[language = Julia]
  function unc_flow(x)
    if x < 0 
      return x - 1
    else 
      return x
    end
  end
\end{lstlisting}

An interval evaluation of the function $\text{unc\_flow}([-1,1])$ would be the union of the output of the different branches, the interval union $[-2, -1) \cup [0, 1]$. For a language like Python, this would require an uncertainty pre-compiler (like that proposed in \cite{gray2019computing}) whereby the python code is first transformed using a pre-compiler, and then evaluated in Python. In Julia, this can be handled natively with its metaprogramming capabilities.

%, and discuss some of the remaining theoretical tasks required to make such a goal a reality.

@@

\section{Probability Boxes}
\label{sec:pboxes}

@@

\section*{Acknowledgements}
We would like to thank our friends and colleagues at the Risk Institute who have supported \texttt{ProbabilityBoundsAnalysis.jl} from its birth, and through their contributions and discussions have made this software possible: Marco De Angelis, Nick Gray, Alex Wimbush and Dominic Calleja. We also thank Dominic, Marco, and Liam Henry's (Clarus Financial Technology, Belfast) detailed review of this paper.
